Bot's Guide to Being Human: Investigating the Effect of Conversational Architecture Elements on the Anthropomorphized Perceptions of Conversational Agents

### 2023-02-02 paper.tex
added notes for more papers (up to [64])

%%\textbf{Content}

%\textit{Affective language}
%* Use words with increased lexical emotional expressiveness by selection from NRC Valence, Arousal and Dominance (NRC-VAD) Lexicon - did not have an effect on engagement. 
%* System with extroverted personality is more irritating, did not like computer using the level of familiar with them and prefer a more reserved system. \cite{andrews2012system}\cmt{[38]}
%* Encouraging personality is rated as easier to use \cite{healey2013relating}\cmt{[39]}

%\textit{Self-Disclosure}
%* High disclosure group had higher enjoyment than the non disclosure group at the end 3 weeks (similar results %around 1 week mark).

%\textit{Humour}
%* Humour did not enhance enjoyment measurement. but qualitative feedback said that humour made the interaction more engaging and immersive

@@

%
%\textbf{Style}

%\textit{Lexical Alignment}
%* Using the words already used by conversational partner resulted in lower perceived cognitive demand, but did not impact perception of habitability (ease of use). \cite{linnemann2018can}\cmt{[15]}
%* Lexical alignment (using same words) resulted in lower cognitive demand, but no impact on annoyance. \cite{huiyang2022improving}\cmt{[17]}
%* Lexical alignment - both words, and structural (grammatical transformations), movie showtimes. There is a decreased workload for substitution and transformation alignment conditions. Also, there is a significant difference in user engagement. \cite{spillner2021talk}\cmt{[18]}

%\textit{Formality}
%* No significant difference between casual and formal conversational styles for enjoyment. \cite{cox2022does}\cmt{[27]}

%\textit{Emoticons}
%* Positive but non-significant trend on helpfulness \cite{wilhelm2022keep}\cmt{[28]}
%* For discussions on mental well beings, agent using emojis is rated higher for enjoyment. However, for agent with discussing physical well being, agent using emojis is rated as lower for enjoyment. \cite{fadhil2018effect}\cmt{[52]}

@@

%
%\textbf{Format}

%\textit{Prosody}
%* More empathetic voice: Enjoyment is marginally significant; high empathy group felt higher ease of use. People expect a high level of agent's social behavior that is matched with their emotional states. \cite{kim2020can}\cmt{[24]}
%* No impact of emotional expressiveness on engagement \cite{zhu2022effects}\cmt{[26]}
%* convenience is higher for default rate \cite{choi2020nobody}\cmt{[54]}

%\textit{Typos}
%* Negatively impacted task attraction. \cite{westerman2019believe}\cmt{[9]}

%\textit{Capitalization}
%* Didn't impact task attraction \cite{westerman2019believe}\cmt{[9]}

@@

%
%\textbf{Style}

%\textit{Lexical Alignment}
%* Using the words already used by conversational partner did not result in higher response accuracy. \cite{linnemann2018can}\cmt{[15]}
%* Lexical alignment (using same words) resulted in higher perceived response accuracy. \cite{huiyang2022improving}\cmt{[17]}

%
%\textbf{Format}

%\textit{Prosody}
%* AM voice - lower speech rate and lower pitch, no significant difference on rating for accuracy. \cite{dubiel2020persuasive}\cmt{[60]}

@@

%\textbf{Content}

%\textit{Affective language}
%* Participants using CA with dynamic, sentiment-adaptive responses reported a higher service encounter satisfaction compared to a CA that provides static, neutral responses \cite{diederich2019emulating}\cmt{[25]}
%* Did not impact user satisfaction between the affective expression or not agent. \cite{yang2017perceived}\cmt{[44]}

%
%\textbf{Style}

%\textit{Disfluency}
%* No significant effects of satisfaction. \cite{pfeifer2009should}\cmt{[12]}

%\textit{Emoticons}
%* Positive but non-significant trend on user satisfaction \cite{wilhelm2022keep}\cmt{[28]}

@@

%
%\textbf{Format}

%\textit{Prosody}
%* Overall satisfaction as well as speech rate satisfaction is higher for the default rate compare to fast rate. Want everyone to understand. Hard to understand the faster rate as communicating with CA over a distance (as oppose to mobile phone. \cite{choi2020nobody}\cmt{[54]}

%\textit{Response Delay}
%* Significant difference in overall user satisfaction, the timing "felt right" and comparing response to human-human communication. \cite{gnewuch2018faster}\cmt{[19]}

@@

%
%\textbf{Style}

%\textit{Lexical Alignment}
%* Lexical alignment - both words, and structural (grammatical transformations), movie showtimes. There is no significant difference in domain knowledge. \cite{spillner2021talk}\cmt{[18]}

%\textit{Disfluency}
%* Participants perceived the agent using fillers like "um" "uh" to be less intelligent. Especially in taks-oriented situations. In social situations, perceive intelligence is slightly increased. \cite{jeong2019exploring}\cmt{[10]}
%* No significant effects of knowledgeable. \cite{pfeifer2009should}\cmt{[12]}

@@

%
%\textbf{Format}

%\textit{Prosody}
%* AM voice - lower speech rate and lower pitch, no significant difference on rating for knowledgeable \cite{dubiel2020persuasive}\cmt{[60]}

@@

%\textbf{Content}

%\textit{Affective Language}
%* Participants rate the agent as lower in competence when it could not feel / express its feelings (patiency). \cite{lee2019s}\cmt{[55]}

%
%\textbf{Style}

%\textit{Formality}
%* No significant difference between casual and formal conversational styles for competence. \cite{cox2022does}\cmt{[27]}

@@

%\textbf{Content}

%\textit{Affective language}
%* Participants using CA with dynamic, sentiment-adaptive responses reported a higher social presence compared to a CA that provides static, neutral responses \cite{diederich2019emulating}\cmt{[25]}
%* Psychological distance: Participants reported they identified with the agent more when the agent was described as being able to feel / express its feelings (patiency). \cite{lee2019s}\cmt{[55]}

%\textit{Humour}
%* Humourous agent was not rated as more social. \cite{ceha2021can}\cmt{[57]}

@@

%\textbf{Format}

%\textit{Prosody}
%* Connectedness higher in the high empathy group. \cite{kim2020can}\cmt{[24]}

%\textit{Typos}
%* Negatively impacted electronic presence \cite{westerman2019believe}\cmt{[9]}

%\textit{Capitalization}
%* Didn't impact electronic presence \cite{westerman2019believe}\cmt{[9]}

%\textit{Response Delay}
%* Significant difference in perceived social presence \cite{gnewuch2018faster}\cmt{[19]}

@@

%\textbf{Content}

%\textit{Affective language}
%* CA with dynamic, sentiment-adaptive responses is perceived as more empathetic than a CA that provides static, neutral responses \cite{diederich2019emulating}\cmt{[25]}
%* Use words with increased lexical emotional expressiveness by selection from NRC Valence, Arousal and Dominance (NRC-VAD) Lexicon - increased the perception of emotional expressiveness of the agent. \cite{zhu2022effects}\cmt{[26]}
%* Agent using affective expressions are seen as more empathetic \cite{yang2017perceived}\cmt{[44]}
%* Empathic chatbot - ask about users feelings and responses in an empathic, supportive manner. There is a statistically significant difference in empathy rating between the advice-only chatbot and the empathic chatbot. \cite{daher2020empathic}\cmt{[58]}

@@

%\textbf{Style}

%\textit{Formality}
%* No significant difference between casual and formal conversational styles for appropriateness of tone. \cite{cox2022does}\cmt{[27]}

%
%\textbf{Format}

%\textit{Prosody}
%* AM voice - lower speech rate and lower pitch, is rated as more persuasive and involved (involved, assertive, active, sincere). There is no difference in powerful (powerful, bold, forceful) \cite{dubiel2020persuasive}\cmt{[60]}

@@

%\textbf{Content}

%\textit{Self-Disclosure}
%* High disclosure group had higher level of intimacy than the non disclosure group at the end 3 weeks (similar results around 1 week mark). \cite{lee2020hear}\cmt{[23]}

%
%\textbf{Style}

%\textit{Lexical Alignment}
%* Using the words already used by conversational partner did not result in higher quality of relationship. \cite{linnemann2018can}\cmt{[15]}

@@

%\textbf{Format}

%\textit{Prosody}
%* Empathetic tone: Intimacy is significantly higher in high empathy group. Similarity is marginally significant. \cite{kim2020can}\cmt{[24]}
%* Default speech rate more intimate than the fast rate, more like a human, more natural and intimate. Fast-rate CA perceived as a mechanical being. \cite{choi2020nobody}\cmt{[54]}

%\textit{Typos}
%* Negatively impacted social attraction. \cite{westerman2019believe}\cmt{[9]}

%\textit{Capitalization}
%* Didn't impact social attraction \cite{westerman2019believe}\cmt{[9]}

@@

%\textbf{Content}

%\textit{Affective language}
%* CA with dynamic, sentiment-adaptive responses is perceived as more human-like than a CA that provides static, neutral responses \cite{diederich2019emulating}\cmt{[25]}
%* Use words with increased lexical emotional expressiveness by selection from NRC Valence, Arousal and Dominance (NRC-VAD) Lexicon - did not have an effect on human-likeness. \cite{zhu2022effects}\cmt{[26]}

@@

%\textbf{Style}

%\textit{Formality}
%* Agents using normal style seemed more humanlike \cite{ouchi2019should}\cmt{[59]}

%\textit{Disfluency}
%* No statistical significance in the perceived difference for perceived humanness overall. Noticed tendency for users to evaluate agent as more human-like in the filler condition. \cite{jeong2019exploring}\cmt{[10]}
%* No significant effects of naturalness. \cite{pfeifer2009should}\cmt{[12]}

@@

%\textbf{Format}

%\textit{Prosody}
%* Expressive prosody did not have an effect on human-likeness. \cite{zhu2022effects}\cmt{[26]}

%\textit{Typos}
%* Made the agents seem less human. \cite{westerman2019believe}\cmt{[9]}

%\textit{Capitalization}
%* Didn't impact humanness \cite{westerman2019believe}\cmt{[9]}

%\textit{Response Delay}
%* Significant difference in perceived humanness \cite{gnewuch2018faster}\cmt{[19]}

@@

%\textbf{Content}

%\textit{Affective Language}
%* System with extroverted personality is less trustworthy \cite{andrews2012system}\cmt{[38]}

%\textit{Self-Disclosure}
%* High disclosure group reported more trust than the non disclosure group. Trust change over time for all groups were not significant. \cite{lee2020hear}\cmt{[23]}

@@

%\textbf{Style}

%\textit{Lexical Alignment}
%* No significant effect on trustworthiness (ability, benevolence and integrity) \cite{linnemann2018can}\cmt{[15]}
%* Lexical alignment (using same words) no significant impact on trustworthiness. \cite{huiyang2022improving}\cmt{[17]}
%* Matched users content (pronoun use, repetition, utterance length) and acoustic variables (speech rate, pitch, loudness) did not significant impact on trustworthiness \cite{hoegen2019end}\cmt{[31]}

%\textit{Disfluency}
%* No significant effects of trust. \cite{pfeifer2009should}\cmt{[12]}

%\textit{Emoticons}
%* No significant impact on trustworthiness. \cite{wilhelm2022keep}\cmt{[28]}

%\textbf{Format}

%\textit{Prosody}
%* AM voice - lower speech rate and lower pitch, no significant difference on rating for truthful. \cite{dubiel2020persuasive}\cmt{[60]}

@@

%\textbf{Content}

%\textit{Affective language}
%* Use words with increased lexical emotional expressiveness by selection from NRC Valence, Arousal and Dominance (NRC-VAD) Lexicon - did not have an effect on likability. \cite{zhu2022effects}\cmt{[26]}
%* Significant difference in the variances of the responses for friendliness of the agent \cite{healey2013relating}\cmt{[39]}
%* Participants rate the agent as lower in warmth when it could not feel / express its feelings (patiency).  \cite{lee2019s}\cmt{[55]}

%\textit{Humour}
%* Humorous agent was not rated as more likeable. But it is rated as having more of a sense of humour. \cite{ceha2021can}\cmt{[57]}

%

@@

%\textbf{Style}

%\textit{Lexical Alignment}
%* Using the words already used by conversational partner did not result in higher likeability. \cite{linnemann2018can}\cmt{[15]}
%* Lexical alignment (using same words) no significant impact on likability. \cite{huiyang2022improving}\cmt{[17]}
%* Lexical alignment - both words, and structural (grammatical transformations), movie showtimes. There is no significant difference in user opinions like friendly, polite or natural. \cite{spillner2021talk}\cmt{[18]}

%\textit{Formality}
%* Casual conversational style perceived warmer than formal style. \cite{cox2022does}\cmt{[27]}
%* Agents using normal style (vs. polite / formal style) seemed more friendly, fun and kind \cite{ouchi2019should}\cmt{[59]}

%\textit{Disfluency}
%* No statistical significance in the perceived difference for likability rating overall. There is significant interaction effect between fillers and context of conversation (task vs. social). Users considered fillers as entertaining and fun in social conversations, but inappropriate in task oriented situations. \cite{jeong2019exploring}\cmt{[10]}
%* No significant effects of likability. \cite{pfeifer2009should}\cmt{[12]}

@@

%\textbf{Format}

%\textit{Prosody}
%*  Emotional expressiveness did not have a significant impact on truthfulness. \cite{dubiel2020persuasive}\cmt{[60]}

@@

@@

@@

* Combined effects - integrated mental model
* Lack of investigations - name a few areas

@@

>>>>

@@

@@

Content: Humour
* No overall significant difference between conventional method vs. agent using humour for enjoyment. Specifically, men enjoyed the agent's jokes more than women.  \cite{miyamoto2017improving}\cmt{[46]}.

Style: Language Formality
* Users spent more time with the Shakespearean chatbot version with more non-task related user inputs. Enjoyment was measured but no report \cite{elsholz2019exploring}\cmt{[61]}

@@

@@

@@

\textbf{Style: Language Formality}
* Modern language had higher usability score compared to Shakespearean language \cite{elsholz2019exploring}\cmt{[61]}
* There is no significant effect on customer satisfaction of the CA between low-status and high-status languages. \cite{habler2019effects}\cmt{[63]}

\textbf{Format: Prosody}
* There is no significant effect on customer satisfaction of the CA between male-gendered and female-gendered voices. \cite{habler2019effects}\cmt{[63]}

@@

@@

@@

\textbf{Style: Language Formality}
*  Low-status language resulted in higher achieved performance ratings \cite{habler2019effects}\cmt{[63]}

\textbf{Format: Prosody}
* There is no significant effect on perceived performance of the CA between male-gendered and female-gendered voices. \cite{habler2019effects}\cmt{[63]}

\textbf{Format: Proactivity}
* Certain proactive styles like notification strategy (conservatively proactive, letting participants know that the agent has found a solution) was evaluated as significantly higher perceived competence compared to the non-strategy. \cite{kraus2020effects}\cmt{[64]}

@@

@@

\textbf{Content: Affective Language (Voice)}
* Model aiming to elicit position emotions from users received higher emotional connection \cite{lubis2019positive}\cmt{[43]}

\textbf{Format: Response Delay}
* Delayed response is associated with higher social presence \cite{gnewuch2022opposing}\cmt{[20]}

\textbf{Content: Affective language and Style: Disfluency}
* Rated higher that agent listened openly to participants' emotions, as that agent can feel waht the user is feeling. \cite{hu2021enhancing}\cmt{[56]}

@@

@@

\textbf{Content: Affective language and Style: Disfluency}
Respond with empathy to the user. \cite{hu2021enhancing}\cmt{[56]}

@@

@@

@@

@@

@@

\textbf{Style: Disfluency}
* (Voice) No significant difference in naturalness rating for About Myself text, but showed that naturalness is higher for fluent system compared to disfluent speech \cite{wester2015artificial}\cmt{[14]}.

@@

@@

\textbf{Format: Prosody}
* No significant difference in trust ratings across different voice prosody settings with differences in pitch. \cite{cambre2019one}\cmt{[62]}

\textbf{Format: Proactivity}
* Overall trust in the system did not differ significantly. The intervention style is less trusted than conservative strategies - too obstructive and imposing. Proactive should be subtle  \cite{kraus2020effects}\cmt{[64]}

@@

@@

@@

\textbf{Content: Humour}
* An agent that uses humourous content is assessed to be more funny \cite{khooshabeh2011does}\cmt{[37]}.

\textbf{Style: Disfluency}
* (Voice) Both About Myself and Speed Dating ratings has shown system scored significantly differently on the Big-5 personality traits. Disfluency makes the voice sound more neurotic, less open, less extrovert and less conscientious \cite{wester2015artificial}\cmt{[14]}.

\textbf{Format: Prosody}
* Low voice pitch rated not as friendly as the other ones. No impact on other traits like assertive and politeness. \cite{cambre2019one}\cmt{[62]}

@@

* Format: Response delay - Novice users perceived the agent as more socially present, but experienced users had a negative effect \cite{gnewuch2022opposing}\cmt{[20]}
* Content: Humour - no overall significant difference between conventional method vs. agent using humour for enjoyment. Specifically, men enjoyed the agent's jokes more than women.  \cite{miyamoto2017improving}\cmt{[46]}.
* Format: Proactivity - significant interaction between proactive dialogue strategies and task difficulty for perceived competence and perceived reliability. For easier tasks, system help is not really required. \cite{kraus2020effects}\cmt{[64]}

@@

\subsection{Relationship between perceptions}

Response delay: social presence has a significant effect on intention to use. 
\cite{gnewuch2022opposing}\cmt{[20]}

Perceived intelligence positively associated with perceived usefulness, perceived ease of use, initial trust, and perceived anthropomorphism. Perceived anthropomorphism is associated with perceived enjoyment, and intention to adopt. Perceived usefulness and perceived enjoyment impacts intention to adopt. \cite{moussawi2021perceptions}\cmt{[36]}

The size of the effect for language we observed is clearly higher than the size of the effect for voice. \cite{habler2019effects}\cmt{[63]}

@@

Combined effects
* integrated mental model \cite{knijnenburg2016inferring}\cmt{[34]}
* Anthropomorphized chatbot (text) using affective language (emotional expressions), self-referencing, emoticons, response delays. Resulted in higher perceived anthropomorphism, which in result led to lower negative word of mouth and loss of trust. anthropomorphic design is beneficial not only to initial trust but also to the ongoing trust relationship \cite{seeger2021chatbots}\cmt{[35]}

Lack of investigations - name a few areas

@@

Evidence for the influence of voice gender and pitch on stereotypical trait attribution. Active trait attribution is missing as most participants attributes to the negative side of traits for CAs. \cite{cambre2019one}\cmt{[62]}
